Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-10-13 manuscript2.RMD
final push

The observations of the present study, from human skeletal muscle, do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Although we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015]. Neither do our findings indicate that insulin *per se*, at least at physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000]. Increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after a single session of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after either 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT. Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].

@@

The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [@macinnis_investigating_2017]. Further, to ensure that the legs were exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days. Every day, the participants showed up in an overnight fasted state and ingested only either protein and glucose or protein and placebo before and during exercise. As evident from the blood data, glucose ingestion thus led to higher levels both of plasma glucose and serum c-peptide compared to placebo, together with the aforementioned standardisations enabling analyses of the effect of ingesting glucose on total- and specific RNA levels and proteins, as well as muscular performance, within-participant. The present study also had a few limitations; firstly, the sample size was smaller than expected and planned for. Initially the minimum limit of 16 participants, according to the a priori power calculation, was met. However, three dropped out during the intervention, possibly leaving our statistical analyses slightly underpowered. Moreover, the intramuscular glycogen stores were not measured, hence it can't be determined whether the glucose ingestion increased intramuscular glucose. Though it may be a reasonable assumption that the glucose ingestion in this design did lead to increased intramuscular glucose and thus energy levels, it cannot be excluded that the participants' intramuscular glycogen stores were topped up from the previous day, and as such, ingesting more glucose had no further benefit. Nevertheless, even if the skeletal muscle of the leg exercising with glucose supplement took up the extra glucose, it did not seem to affect any of our main outcome measures.

>>>>

The observations of the present study from human skeletal muscle do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Although we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015]. Neither do our findings indicate that insulin *per se*, at least at physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000]. Increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after a single session of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT. Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].

@@

The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [@macinnis_investigating_2017]. Further, to ensure that the legs were exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days[@halperin_threats_2015]. Every day, the participants showed up in an overnight fasted state and ingested only either protein and glucose or protein and placebo before and during exercise. As evident from the blood data, glucose ingestion thus led to higher levels both of plasma glucose and serum c-peptide compared to placebo, together with the aforementioned standardisations enabling analyses of the effect of ingesting glucose on total- and specific RNA levels and proteins, as well as muscular performance, within-participant. The present study also had a few limitations; firstly, the sample size was smaller than expected and planned for. Initially the minimum limit of 16 participants, according to the a priori power calculation, was met. However, three dropped out during the intervention, possibly leaving our statistical analyses slightly underpowered. Moreover, the intramuscular glycogen stores were not measured, hence it can't be determined whether the glucose ingestion increased intramuscular glucose. Though it may be a reasonable assumption that the glucose ingestion in this design did lead to increased intramuscular glucose and thus energy levels, it cannot be excluded that the participants' intramuscular glycogen stores were topped up from the previous day, and as such, ingesting more glucose had no further benefit. Nevertheless, even if the skeletal muscle of the leg exercising with glucose supplement took up the extra glucose, it did not seem to affect any of our main outcome measures.
